---
output:
  md_document:
    variant: markdown_github
---
<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r Setup, include=FALSE}
library(rroad)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "vignettes/README-"
)
set.seed(1)
```

# rroad R package
Set of tools for monitoring road condition.

Currently, the following features are supported:

  - International Roughness Index (IRI) computation
    - continuously increasing segment
    - fixed length overlapping segments with an offset

[![Build Status](https://travis-ci.org/vsimko/rroad.svg)](https://travis-ci.org/vsimko/rroad)
[![codecov.io](https://codecov.io/github/vsimko/rroad/coverage.svg?branch=master)](https://codecov.io/github/vsimko/rroad?branch=master)
[![CRAN Status](http://www.r-pkg.org/badges/version/rroad)](http://cran.r-project.org/package=rroad)
[![Rdoc](http://www.rdocumentation.org/badges/version/rroad)](http://www.rdocumentation.org/packages/rroad)
[![Dependency Status](https://dependencyci.com/github/vsimko/rroad/badge)](https://dependencyci.com/github/vsimko/rroad)

## Download and Install

To download the development version of the package, type the following at the R command line:
```{r Install latest, eval=FALSE}
install.packages("devtools")
devtools::install_github("vsimko/rroad")
```

To download the release version of the package on CRAN, type the following at the R command line:
```{r Install from CRAN, eval=FALSE}
install.packages("rroad")
```


## Examples

```{r iri_from_random_profile, fig.width=8, fig.height=4}
profile <- rnorm(10000)
iri <- CalculateIRIperSegments(profile, IRI_COEF_100, 20)
par(mfrow = c(1,2)) # space for two diagrams
plot(profile, type = "l",
  xlab = "Distance [dm]", ylab = "Profile [mm]",
  main = "Read profile (Laser measurement)")
plot(iri, type = "s",
  xlab = "Segment", ylab = "IRI [m/km]",
  main = "International Roughness Index (IRI)\nsample = 10cm, segment = 20m")
```

## References

<a id="Sayers1986"/>
Sayers, Michael W; Gillespie, Thomas D; Queiroz, Cesar A.V. 1986
**The International Road Roughness Experiment (IRRE) : establishing correlation and a calibration standard for measurements**.
World Bank technical paper ; no. WTP 45. Washington, DC : The World Bank. [Link]

[Link]: http://documents.worldbank.org/curated/en/326081468740204115/The-International-Road-Roughness-Experiment-IRRE-establishing-correlation-and-a-calibration-standard-for-measurements
