---
output:
  md_document:
    variant: markdown_github
---
<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r Setup, include=FALSE}
library(rroad)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "inst/webimg/README-"
)
```

# biwavelet R package
Set of tools for monitoring road condition

- International Roughness Index (IRI) computation

[![Build Status](https://travis-ci.org/vsimko/rroad.svg)](https://travis-ci.org/vsimko/rroad)
[![codecov.io](https://codecov.io/github/vsimko/rroad/coverage.svg?branch=master)](https://codecov.io/github/vsimko/rroad?branch=master)

## Download and Install

To download the development version of the package, type the following at the R command line:
```{r Install latest, eval=FALSE}
install.packages("devtools")
devtools::install_github("vsimko/rroad")
```

## Examples

```{r iri_from_random_profile, fig.width=8, fig.height=4}
profile <- rnorm(10000)
iri <- CalculateIRI(profile, IRI_COEF_100, 20)
par(mfrow = c(1,2)) # space for two diagrams
plot(profile, type="l",
  xlab="Distance [cm]", ylab="Profile [mm]",
  main="Read profile (Laser measurement)")
plot(iri, type="s",
  xlab="Segment", ylab="IRI [m/km]",
  main="International Roughness Index (IRI)\nsample = 10cm, segment = 20m")
```

## References

<a id="Sayers1986"/>
Sayers, M.W., Gillespie, T.D. & Queiroz, C. a V., 1986.
**The International Road Roughness Experiment - Establishing Correlation and a Calibration Standard for Measurements**.
Washington, D.C.: Transportation Department.
